---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
author: "GenDoc94"
date: "2024-08-04"
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

1.  Load the data
```{r reading_data, results='hide'}
linkURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(linkURL, destfile = "activity.zip")
unzip(zipfile = "activity.zip")
activity <- read.csv("activity.csv", header = TRUE, sep = ",")
file.remove("activity.zip")
```

2.  Transform the data
```{r transforming_data, results='hide'}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

3.  See the head of the data
```{r head_data}
head(activity)
str(activity)
```

## What is mean total number of steps taken per day?

1.  Calculating the steps by day
```{r calculate_steps}
day_steps <- aggregate(steps ~ date, activity, FUN = sum)
```

2.  Histogram of number of steps
```{r steps_histogram}
g <- ggplot(day_steps, aes(x=steps))
g + 
        geom_histogram(fill = "darkblue", binwidth = 600) +
        labs(title = "Histogram of Steps by day", x = "Steps", y = "Freq") +
        theme_bw()
```

3.  Calculate the mean and median of total steps per day
```{r calculate_mnmd, results='hold'}
step_mean <- mean(day_steps$steps, na.rm = TRUE)
step_median <- median(day_steps$steps, na.rm = TRUE)
```
The mean is `r step_mean` and the median is `r step_median`.

## What is the average daily activity pattern?

1.  Calculating interval
```{r interval}
intervaldb <- group_by(activity, interval) %>% 
                summarise(steps=mean(steps, na.rm = TRUE)) %>%
                unique()
```

2.  Plotting the interval
```{r plot_interval}
g <- ggplot(intervaldb, aes(x=interval, y=steps))
g + 
        geom_line(color = "darkblue", linewidth = 0.5) +
        labs(title = "Average of daily steps", x = "Interval", y = "Average steps/day") +
        theme_bw()
```

3.  5-minute interval with the maxium number of steps
```{r max_interval}
max_interval <- intervaldb[which.max(intervaldb$steps),1]
```
The maximum interval is `r max_interval`.

## Imputing missing values

1.  Calculating total missing values in the dataset
```{r total_missing}
nas <- sum(is.na(activity$steps))
```
Total number of missing values: `r nas`.

2. Imputing data on missing values. I used mean values of steps for filling the missing values. Firstly matched by interval across dates.


## Are there differences in activity patterns between weekdays and weekends?

